Analytic Occultation Light Curves in Reflected Light

### 2020-04-06 tex/ms.tex
omg so much math

The key to identifying the case corresponding to a given configuration is
to determine whether or not the limb of the occultor intersects the terminator
of the body, and if so, the points of intersection.
While we perform the integration in frame $\mathcal{F}'$, finding the points
of intersection with the terminator is easier if we temporarily switch to
the frame $\mathcal{F}''$ (via Equation~\ref{eq:R''}) in which the terminator
is parallel to the $x''$-axis.
In this frame, the equations defining the terminator and the limb of the
occultor are, respectively,
%
\begin{proof}{}
    y''_1(x'') & = b \sqrt{1 - x''^2}
    \nonumber                                               \\
    y''_2(x'') & = y''_o \pm \sqrt{r_o^2 - (x'' - x''_o)^2}
\end{proof}
%
where
%
\begin{proof}{}
    x''_o = b_o\sin\theta'
    \nonumber \\
    y''_o = b_o\cos\theta'
\end{proof}
%
are the coordinates of the occultor in $\mathcal{F}''$.
%
We wish to find the set of $N$ points
$\{x_n''~\big|~0~\le~n~<~N\}$
for which
$y''_1(x_n'') - y''_2(x_n'') = 0$. Following \citet{Luger2017}, we may
express this condition as the quartic equation
%
\begin{proof}{}
    \label{eq:quartic}
    A {x''}^4 + B {x''}^3 + C {x''}^2 + D {x''} + E = 0
\end{proof}
%
with coefficients
%
\begin{proof}{}
    \label{eq:quartic-coeffs}
    A &= (1 - b^2)^2
    \nonumber \\
    B &= -4 x''_o (1 - b^2)
    \nonumber \\
    C &= -2 \bigg(
    b^4
    + r_o^2
    - 3 {x''_o}^2
    - {y''_o}^2
    - b^2 \big(1 + r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    \bigg)
    \nonumber \\
    D &= -4 x''_o (b^2 - r_o^2 + {x''_o}^2 + {y''_o}^2)
    \nonumber \\
    E &=
    b^4
    - 2 b^2 \big(r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    + \big(r_o^2 - {x''_o}^2 - {y''_o}^2\big)^2
    \quad.
\end{proof}
%
Although closed-form solutions to quartic equations exist
\citep[see, e.g.,][who solve for the area of overlap between two ellipses
    analytically]{Hughes2011}, they are prone to significant numerical
instabilities. Instead, we solve for the roots of the quartic
numerically by casting it
as an eigenvalue problem \citep[e.g.,][]{Edelman1995} and polish the
results with a few iterations of Newton's method. We find that this is
reasonably computationally efficient and
yields roots with precision within a couple orders of magnitude of machine
epsilon (see \S\ref{sec:performance}).

@@

Cases 6--10 (see Figure~\ref{fig:cases}) correspond to configurations in which
the limb of the occultor intersects with the terminator at either one point
(case 6) or two points (cases 7--10). Because of these intersections, we
cannot simply re-weight the emitted light solution, as the integration
boundaries are now different. In general, we may compute the flux by
integrating the components of the Green's basis $\bg$
over the region $S$ bounded by three curves, which we denote
$P$, $T$, and $Q$. These are shown in orange
in Figure~\ref{fig:cases} and presented in more detail in
Figure~\ref{fig:geometry}.
Curve $P$ is a segment of the limb of
the occultor, parametrized by the angle $\phi \in [\phi_1, \phi_2]$;
curve $T$ is a segment of the terminator,
parametrized by the angle $\xi \in [\xi_1, \xi_2]$;
and curve $Q$ is a segment of the limb of the occulted body,
parametrized by the angle $\lambda \in [\lambda_1, \lambda_2]$.
The endpoints $\phi_1, \phi_2, \xi_1, \xi_2, \lambda_1,$ and
$\lambda_2$ are functions of the solutions to the quartic from
\S\ref{sec:which-case} and will be presented in \S\ref{sec:sT}.

@@

\subsection{Computing the integrals $\sT$}
\label{sec:sT}
%
In the previous sections, we discussed how to identify the case
corresponding to a specific configuration of the occultor and the
illumination source. We showed how in some cases
(1--5; \S\ref{sec:cases-easy}) the total flux may be
computed by exploiting the classical \starry integrals (Equation~\ref{eq:fI}).
In all other cases (6--14; \S\ref{sec:cases-hard} and
\S\ref{sec:cases-pathological}), however, the flux computation involves
evaluation of
Equation~(\ref{eq:fS}), where the solution vector $\mathfrak{s}^\top$
is the vector of integrals of each of the terms in the Green's basis $\bg$
over a region $S$ of the
projected disk of the occultor (Equation~\ref{eq:sTint}).
%
As in \citet{Luger2019}, the approach to computing $\sT$
is to use Green's theorem to transform the surface integrals into line
integrals along the curves $P$, $T$, and $Q$
(see Figure~\ref{fig:geometry}). Specifically, we write%
\footnote{%
    In this section, we deliberately drop the prime symbols from $x'$ and
    $y'$ for clarity. We also drop the dependence of $\sT$ and the primitive
    integrals on the geometrical parameters $b, \theta', b_o, r_o$.
}
%
\begin{align}
    \label{eq:greens}
    \sT
     & =
    \iint\limits_{S}
    \bg^\top(x, y)
    \ \dd x \ \dd y
    \nonumber \\[0.5em]
     & =
    \oint \bvec{G} (x, y) \cdot
    \dd \bvec{r} (x, y)
    \quad,
\end{align}
%
where $\bvec{G} (x, y)$
is an array of two-dimensional Cartesian vectors chosen such that its
exterior derivative is $\bg$,
%
\begin{align}
    \label{eq:DGg}
    \frac{\dd \bvec{G}_{y}(x, y)}{\dd x}
    - \frac{\dd \bvec{G}_{x}(x, y)}{\dd y} = \bg(x, y)
    \quad,
\end{align}
%
and
%
\begin{align}
    \dd \BF{r} (x, y) & =
    \left(\frac{\dd x}{\dd \varphi}\right) \dd \varphi \, \xhat +
    \left(\frac{\dd y}{\dd \varphi}\right) \dd \varphi \, \yhat
    \quad,
\end{align}
%
where $\varphi$ is the parametrized angle along the integration path.
%
\citet{Luger2019} showed that one possible solution to
Equation~(\ref{eq:DGg}) consists of the vector whose $n^\mathrm{th}$
component is given by
%
\begin{align}
    \label{eq:G}
    \BF{G}_n (x, y) & =
    \begin{dcases}
        %
        x^{\frac{\mu + 2}{2}}
        y^{\frac{\nu}{2}}
        \,\yhat
         & \qquad \mu, \nu \, \mathrm{even}
        \\[1em]
        %
        \frac{1-z(x, y)^3}{3(1-z(x, y)^2)}\bigg(-y \, \xhat + x \, \yhat\bigg)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        x^{l-2}
        z(x, y)^3
        \,\xhat
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{even}
        \\[1em]
        %
        x^{l-3}
        y
        z(x, y)^3
        \,\xhat
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{odd}
        \\[1em]
        %
        x^{\frac{\mu-3}{2}}
        y^{\frac{\nu-1}{2}}
        z(x, y)^3
        \,\yhat
         & \qquad \mathrm{otherwise,}
    \end{dcases}
\end{align}
%
where the indices $l, m, \mu, \nu$ are given by Equation~(\ref{eq:l-m-mu-nu}).

We showed in the previous sections that there are at most three curves
$P$, $T$, and $Q$ bounding a given closed surface of integration
(see Figure~\ref{fig:geometry}). We may therefore express
Equation~(\ref{eq:greens}) as
%
\begin{align}
    \label{eq:sT}
    \sT & =
    \mathfrak{p}^\top + \mathfrak{t}^\top + \mathfrak{q}^\top
    \quad,
\end{align}
%
where we define the primitive integrals
%
\begin{align}
    \label{eq:pT}
    \mathfrak{p}^\top
     & =
    \int\limits_{\phi_1}^{\phi_2}
    \mathbf{G}(x_p, y_p)
    \cdot \dd \mathbf{r}(x_p, y_p)
    %
    \\
    %
    \label{eq:tT}
    \mathfrak{t}^\top
     & =
    \int\limits_{\xi_1}^{\xi_2}
    \mathbf{G}(x_t, y_t)
    \cdot \dd \mathbf{r}(x_t, y_t)
    %
    \\
    %
    \label{eq:qT}
    \mathfrak{q}^\top
     & =
    \int\limits_{\lambda_1}^{\lambda_2}
    \mathbf{G}(x_q, y_q)
    \cdot \dd \mathbf{r}(x_q, y_q)
\end{align}
%
to be the line integrals of $\mathbf{G}$ along each of the curves
$P$, $T$, and $Q$, respectively, where the coordinates are parametrized
in terms of $\varphi$ as follows:
%
\\[1em]
%
\begin{minipage}{0.31\linewidth}
    \begin{align}
        x_p & = r_o \cos\varphi
        \nonumber                     \\
        y_p & = b_o + r_o \sin\varphi
        \nonumber
    \end{align}
\end{minipage}
%
\begin{minipage}{0.35\linewidth}
    \begin{align}
        x_t & = \cos\theta' \cos\varphi - b \sin\theta' \sin\varphi
        \nonumber                                                   \\
        y_t & = \sin\theta' \cos\varphi + b \cos\theta' \sin\varphi
        \nonumber
    \end{align}
\end{minipage}
%
\begin{minipage}{0.31\linewidth}
    \begin{align}
        \label{eq:xy_pqt}
        x_q & =\cos\varphi
        \nonumber           \\
        y_q & = \sin\varphi
        \quad.
    \end{align}
\end{minipage}
%
\\[1em]

\xxx{We perform the integration in a frame $\mathcal{F}'$
    in which the occultor is aligned with the $+y'$-axis,
    in a counter-clockwise direction relative to the center
    of the integration region $S$.}

>>>>

The key to identifying the case corresponding to a given configuration is
to determine whether or not the limb of the occultor intersects the terminator
of the body, and if so, the points of intersection.
While we perform the integration in frame $\mathcal{F}'$, finding the points
of intersection with the terminator is easier if we temporarily switch to
the frame $\mathcal{F}''$ (via Equation~\ref{eq:R''}) in which the terminator
is parallel to the $x''$-axis.
In this frame, the equations defining the terminator and the limb of the
occultor are, respectively,
%
\begin{proof}{}
    y''_1(x'') & = b \sqrt{1 - x''^2}
    \nonumber                                               \\
    y''_2(x'') & = y''_o \pm \sqrt{r_o^2 - (x'' - x''_o)^2}
\end{proof}
%
where
%
\begin{proof}{}
    x''_o = b_o\sin\theta'
    \nonumber \\
    y''_o = b_o\cos\theta'
\end{proof}
%
are the coordinates of the occultor in $\mathcal{F}''$.
%
We wish to find the set of $N$ points
$\BS{x''} = \{x_n''~\big|~0~\le~n~<~N\}$
for which
$y''_1(x_n'') - y''_2(x_n'') = 0$. Following \citet{Luger2017}, we may
express this condition as the quartic equation
%
\begin{proof}{}
    \label{eq:quartic}
    A {x''}^4 + B {x''}^3 + C {x''}^2 + D {x''} + E = 0
\end{proof}
%
with coefficients
%
\begin{proof}{}
    \label{eq:quartic-coeffs}
    A &= (1 - b^2)^2
    \nonumber \\
    B &= -4 x''_o (1 - b^2)
    \nonumber \\
    C &= -2 \bigg(
    b^4
    + r_o^2
    - 3 {x''_o}^2
    - {y''_o}^2
    - b^2 \big(1 + r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    \bigg)
    \nonumber \\
    D &= -4 x''_o (b^2 - r_o^2 + {x''_o}^2 + {y''_o}^2)
    \nonumber \\
    E &=
    b^4
    - 2 b^2 \big(r_o^2 - {x''_o}^2 + {y''_o}^2\big)
    + \big(r_o^2 - {x''_o}^2 - {y''_o}^2\big)^2
    \quad.
\end{proof}
%
Although closed-form solutions to quartic equations exist
\citep[see, e.g.,][who solve for the area of overlap between two ellipses
    analytically]{Hughes2011}, they are prone to significant numerical
instabilities. Instead, we solve for the roots of the quartic
numerically by casting it
as an eigenvalue problem \citep[e.g.,][]{Edelman1995} and polish the
results with a few iterations of Newton's method. We find that this is
reasonably computationally efficient and
yields roots with precision within a couple orders of magnitude of machine
epsilon (see \S\ref{sec:performance}).

@@

Cases 6--10 (see Figure~\ref{fig:cases}) correspond to configurations in which
the limb of the occultor intersects with the terminator at either one point
(case 6) or two points (cases 7--10). Because of these intersections, we
cannot simply re-weight the emitted light solution, as the integration
boundaries are now different. In general, we may compute the flux by
integrating the components of the Green's basis $\bg$
over the region $S$ bounded by three curves, which we denote
$P$, $T$, and $Q$. These are shown in orange
in Figure~\ref{fig:cases} and presented in more detail in
Figure~\ref{fig:geometry}.
Curve $P$ is a segment of the limb of
the occultor, parametrized by the angle $\phi \in [\phi_0, \phi_1]$;
curve $T$ is a segment of the terminator,
parametrized by the angle $\xi \in [\xi_0, \xi_1]$;
and curve $Q$ is a segment of the limb of the occulted body,
parametrized by the angle $\lambda \in [\lambda_0, \lambda_1]$.
The endpoints $\phi_0, \phi_1, \xi_0, \xi_1, \lambda_0,$ and
$\lambda_1$ are functions of the solutions to the quartic from
\S\ref{sec:which-case} and will be presented in \S\ref{sec:sT}.

@@

\subsection{Computing the integrals $\sT$}
\label{sec:sT}
%
In the previous sections, we discussed how to identify the case
corresponding to a specific configuration of the occultor and the
illumination source. We showed how in some cases
(1--5; \S\ref{sec:cases-easy}) the total flux may be
computed by exploiting the classical \starry integrals (Equation~\ref{eq:fI}).
In all other cases (6--14; \S\ref{sec:cases-hard} and
\S\ref{sec:cases-pathological}), however, the flux computation involves
evaluation of
Equation~(\ref{eq:fS}), where the solution vector $\mathfrak{s}^\top$
is the vector of integrals (in the frame $\mathcal{F}'$)
of each of the terms in the Green's basis $\bg$
over a region $S$ of the
projected disk of the occultor (Equation~\ref{eq:sTint}).
%
As in \citet{Luger2019}, the approach to computing $\sT$
is to use Green's theorem to transform the surface integrals into line
integrals along the curves $P$, $T$, and $Q$
(see Figure~\ref{fig:geometry}). Specifically, we write%
\footnote{%
    In this section, we deliberately drop the dependence of $\sT$ and
    the primitive
    integrals on the geometrical parameters $b, \theta', b_o, r_o$
    for clarity.
}
%
\begin{align}
    \label{eq:greens}
    \sT
     & =
    \iint\limits_{S}
    \bg^\top(x', y')
    \ \dd x' \ \dd y'
    \nonumber \\[0.5em]
     & =
    \oint \bvec{G}^\top (x', y') \cdot
    \dd \bvec{r} (x', y')
    \quad,
\end{align}
%
where $\bvec{G} (x', y')$
is an array of two-dimensional Cartesian vectors chosen such that its
exterior derivative is $\bg$,
%
\begin{align}
    \label{eq:DGg}
    \frac{\dd \bvec{G}_{y'}(x', y')}{\dd x'}
    - \frac{\dd \bvec{G}_{x'}(x', y')}{\dd y'} = \bg(x', y')
    \quad,
\end{align}
%
and
%
\begin{align}
    \dd \BF{r} (x', y') & =
    \left(\frac{\dd x'}{\dd \varphi}\right) \dd \varphi \, \xhat' +
    \left(\frac{\dd y'}{\dd \varphi}\right) \dd \varphi \, \yhat'
    \quad,
\end{align}
%
where $\varphi$ is the parametrized angle along the integration path
and the integral is taken in a counter-clockwise direction relative to
the center of the integration region.
%
\citet{Luger2019} showed that one possible solution to
Equation~(\ref{eq:DGg}) consists of the vector whose $n^\mathrm{th}$
component is given by
%
\begin{align}
    \label{eq:G}
    \BF{G}_n (x, y) & =
    \begin{dcases}
        %
        x'^{\frac{\mu + 2}{2}}
        y'^{\frac{\nu}{2}}
        \,\yhat'
         & \qquad \mu, \nu \, \mathrm{even}
        \\[1em]
        %
        \frac{1-z(x', y')^3}{3(1-z(x', y')^2)}\bigg(-y' \, \xhat' + x' \, \yhat'\bigg)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        x'^{l-2}
        z(x', y')^3
        \,\xhat
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{even}
        \\[1em]
        %
        x'^{l-3}
        y'
        z(x', y')^3
        \,\xhat
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{odd}
        \\[1em]
        %
        x'^{\frac{\mu-3}{2}}
        y'^{\frac{\nu-1}{2}}
        z(x', y')^3
        \,\yhat
         & \qquad \mathrm{otherwise,}
    \end{dcases}
\end{align}
%
where the indices $l, m, \mu, \nu$ are given by Equation~(\ref{eq:l-m-mu-nu}).

We showed in the previous sections that there are at most three curves
$P$, $T$, and $Q$ bounding a given closed surface of integration
(see Figure~\ref{fig:geometry}). We may therefore express
Equation~(\ref{eq:greens}) as
%
\begin{proof}{}
    \label{eq:sT}
    \sT & =
    \mathfrak{p}^\top + \mathfrak{t}^\top + \mathfrak{q}^\top
    \quad,
\end{proof}
%
where we define the primitive integrals%
\footnote{%
    The components of the vectors $\mathfrak{p}^\top$ and
    $\mathfrak{q}^\top$ are analogous to the primitive integrals
    $\mathcal{P}$ and $\mathcal{Q}$
    defined in Equations~(30)--(32) in \citet{Luger2019}, although the
    integration limits of both and the sense of integration of $\mathcal{P}$
    are different.
}
%
\begin{proof}{}
    \label{eq:pT}
    \mathfrak{p}^\top
    & =
    \int\limits_{\BS{\phi}}
    \mathbf{G}^\top(x'_p, y'_p)
    \cdot \dd \mathbf{r}(x'_p, y'_p)
    %
    \\
    %
    \label{eq:tT}
    \mathfrak{t}^\top
    & =
    \int\limits_{\BS{\xi}}
    \mathbf{G}^\top(x'_t, y'_t)
    \cdot \dd \mathbf{r}(x'_t, y'_t)
    %
    \\
    %
    \label{eq:qT}
    \mathfrak{q}^\top
    & =
    \int\limits_{\BS{\lambda}}
    \mathbf{G}^\top(x'_q, y'_q)
    \cdot \dd \mathbf{r}(x'_q, y'_q)
\end{proof}
%
to be the line integrals of $\mathbf{G}$ along each of the curves
$P$, $T$, and $Q$, respectively, where the coordinates along each
curve are parametrized
in terms of $\varphi$ as follows:
%
\\[1em]
%
\begin{minipage}{0.31\linewidth}
    \begin{align}
        x'_p & = r_o \cos\varphi
        \nonumber                      \\
        y'_p & = b_o + r_o \sin\varphi
        \nonumber
    \end{align}
\end{minipage}
%
\begin{minipage}{0.35\linewidth}
    \begin{align}
        x'_t & = \cos\theta' \cos\varphi - b \sin\theta' \sin\varphi
        \nonumber                                                    \\
        y'_t & = \sin\theta' \cos\varphi + b \cos\theta' \sin\varphi
        \nonumber
    \end{align}
\end{minipage}
%
\begin{minipage}{0.31\linewidth}
    \begin{proof}{}
        \label{eq:xy_pqt}
        x'_q & =\cos\varphi
        \nonumber           \\
        y'_q & = \sin\varphi
        \quad,
    \end{proof}
\end{minipage}
%
\\[1em]
%
and we define
%
\begin{align}
    \int\limits_{\BS{\varphi}} & \equiv
    %
    \int\limits_{\varphi_{0}}^{\varphi_{1}}
    +
    \int\limits_{\varphi_{2}}^{\varphi_{3}}
    +
    \cdots
    +
    \int\limits_{\varphi_{N - 2}}^{\varphi_{N - 1}}
    %
    \nonumber                           \\
                               & \equiv
    \sum_{i = 0}^{\frac{N - 1}{2}}
    \int\limits_{\varphi_{2i}}^{\varphi_{2i+1}}
\end{align}
%
to be the sum of definite integrals between pairs of limits $\varphi_i$
arranged in an array $\BS{\varphi}$ of length $N$.
For cases involving integration over a single closed region, this
corresponds to a single integral. For cases 11--14, this corresponds to
the sum of two integrals, one for each of the two disjoint regions.
In the next three sections, we derive the solutions to each of these
primitive integrals.

\subsection{The integral along the occultor limb, $\mathfrak{p}^\top$}
\label{sec:pT}
%
In this section we present a solution to Equation~(\ref{eq:pT}). The first
order of business is to derive expressions for the integration limits
$\BS{\phi}$. Depending on the integration case, these limits will correspond
to the point of intersection between the limb of the occultor and the
limb of the occulted body and/or the point of intersection between the limb
of the occultor and the terminator of the occulted body.
The former is given by
\citep[c.f. Equation~24 in][]{Luger2019}
%
\begin{align}
    \phi_l & = \frac{\pi}{2} \pm \left(\arcsin\left(\frac{1 - r_o ^ 2 - b_o ^ 2}{2 b_o r_o}\right) - \frac{\pi}{2}\right)
    \quad,
\end{align}
%
where the sign is chosen such that the point
$(r_o\cos\phi_l, b_o + r_o\sin\phi_l)$ is on the dayside of the occulted body,
and the latter (of which there may be multiple) is given by
%
% Introduce the function
% %
% \begin{proof}{}
%     \textrm{\faAdjust}(x, y) &=
%     \begin{cases}
%         \textrm{\faCircleO} & \qquad -x \sin\theta' + y\cos\theta' \ge b \sqrt{1 - (x\cos\theta' + y\sin\theta') ^ 2}
%         \\
%         \textrm{\faCircle}  & \qquad \mathrm{otherwise}
%     \end{cases}
% \end{proof}
% %
% which returns \faCircleO\ if point $(x, y)$ is on the dayside of the body and
% \faCircle\ if it is on the nightside.
%
%
\begin{align}
    \BS{\phi_t} & =
    \theta' + \mathrm{arctan2}\left(b\sqrt{1 - \BS{x''}^2 - y''_o}, \BS{x''} - x''_o\right)
    \quad
\end{align}
%
where $\BS{x''}$ are the roots of the quartic (Equation~\ref{eq:quartic}).
These angles are then sorted and placed in the array $\BS{\phi}$ such that
the integration is always performed in a counter-clockwise sense about
the center of the integration region.
%
The left panel in Figure~\ref{fig:geometry} shows a configuration in which
the lower integration limit $\phi_0 = 158.2^\circ$ corresponds to the
point of intersection between the limbs of the two bodies and the upper
integration limit $\phi_1 = 223.3^\circ$ corresponds to the
limb-terminator intersection. Both angles are measured counter-clockwise
from the $x'$-axis.

\xxx{OK, now solve the integral.}

Let
%
\begin{align}
    \BS{\kappa} = \BS{\phi} + \frac{\pi}{2}
\end{align}
